---
title: "optimization_test"
author: "Zijun Lu"
date: "October 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Testing some codes to see the performance for optimization. 


```{r}

setwd("~/Dropbox/lammps/PMMA_big/atom600")
source("https://raw.githubusercontent.com/edwardcooper/mlmodel_select/master/timeRecord_functions.R")

MSD=function(data){
  (data$xu-data$xu[1])^2+ (data$yu-data$yu[1])^2+ (data$zu-data$zu[1])^2
}
library(foreach)
library(magrittr)
library(microbenchmark)






timeRecordB()
### Load the data
atom.600.1=read.table(file="atom.600_1.txt",header=FALSE, sep=" ", stringsAsFactors = FALSE, fill=TRUE
                      ,col.names = c("atom-id","type","mol","xu","yu","zu")
                      ,colClasses = c("numeric","numeric","numeric","numeric","numeric","numeric")
                      ,na.strings = c("ITEM:","TIMESTEP","NUMBER","OF","ATOMS","BOX","BOUNDS", "pp","id","type","mol","xu","yu","zu")
)
timeRecordB(output_message = "Load in data")
# Clear out all rows with NAs.
atom.600.1=atom.600.1[complete.cases(atom.600.1),]
gc()
```

It takes about 9.17 mins to load data using read.table.

## try to use fread instead of read.table to read file into R.

```{r}
setwd("~/Dropbox/lammps/PMMA_big/atom600")
timeRecordB()
library(data.table)
### Load the data with fread
atom.600.1_fread=fread(input="atom.600_1.txt", sep=" ", stringsAsFactors = FALSE, fill=TRUE
                      #,col.names = c("atom-id","type","mol","xu","yu","zu")
                      ,colClasses = c("numeric","numeric","numeric","numeric","numeric","numeric","character","character")
                      ,na.strings = c("ITEM:","TIMESTEP","NUMBER","OF","ATOMS","BOX","BOUNDS", "pp","id","type","mol","xu","yu","zu")
                      
)
timeRecordB(output_message = "Load in data fread")
# select the non-NA columns
atom.600.1_fread=atom.600.1_fread[,.(V1,V2,V3,V4,V5,V6)]
# clear the rows that have NA values, V6 contains the most NAs. 
atom.600.1_fread=atom.600.1_fread[complete.cases(atom.600.1_fread[,V6])]
colnames(atom.600.1_fread)=c("atom.id","type","mol","xu","yu","zu")
head(atom.600.1_fread,10)
```

It takes 2 mins 45 seconds to load data with fread. 


```{r}
setwd("~/Dropbox/lammps/PMMA_big/atom600")
timeRecordR(unit="min")%>%select(output_message,run_time)%>%filter(output_message!="None")
```

The clear winner is using fread instead of read.table. About 3 times faster. 

```{r}

library(microbenchmark)

microbenchmark(dplyr=atom.600.1%>%filter(atom.id==1)%>%select(xu,yu,zu)%>%MSD
               ,base_use=atom.600.1[atom.600.1[,"atom.id"]==1,]%>%.[,c("xu","yu","zu")]%>%MSD
               ,mix_use=atom.600.1[atom.600.1[,"atom.id"]==1,]%>%select(xu,yu,zu)%>%MSD
               )
```

Use the first expression from dplyr pacakge, which is the fastest. 



```{r}

atom.600.1[,"atom.id"]%>%max()
atom.600.1[,"atom.id"]%>%min()
atom.600.1[,"mol"]%>%max()
atom.600.1[,"mol"]%>%min()
```
 
This is faster than select. 


```{r}
microbenchmark(
  
  MSD.all.matrix.600=foreach(n=1:2,.combine=rbind)%do%{ 
    atom.600.1%>%filter(atom.id==n)%>%select(xu,yu,zu)%>%MSD
  } 
  ,
  MSD.all.matrix.600.2=foreach(n=1:2,.combine=rbind)%do%{ 
    atom.600.1%>%filter(atom.id==n)%>%select(xu,yu,zu)
  }
  ,times = 10
)
```

These two expressions use about the same amount of time. 

The performance bottleneck is not calculating the MSD but searching for all atom.id==1 and select xu,yu,zu values to calculate MSD 



## Try to use data.table instead of dplyr 

```{r}
library(data.table)
atom.600.1.data_table=atom.600.1%>%as.data.table
timeRecordB()
microbenchmark(
  
  MSD.all.matrix.600=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1%>%filter(atom.id==n)%>%select(xu,yu,zu)%>%MSD
  } 
  ,
  MSD.all.matrix.600.2=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1%>%filter(atom.id==n)%>%select(xu,yu,zu)
  }
  ,
  MSD.all.matrix.600.3=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1.data_table%>%.[atom.id==n]%>%.[,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.4=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,times = 10
)
timeRecordB(output_message = "benchmark data.table searching")
```


Using data.table gives a huge speed-up but the exact speed-up is hard to say. Becasue the maximum and minimum fluctuates a lot. 

```{r}
timeRecordB()
microbenchmark(
  MSD.all.matrix.600.3=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n]%>%.[,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.4=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,times = 200
)
```

Compare using onestep and two step calculation. The result is clear that the one step search is faster than a two step search. 

Later calculation on the entire dataset confirms the faster in speed in about 5 seconds in the total calculation time of around 300 seconds. 



```{r}
microbenchmark(
   MSD.all.matrix.600.1=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1.data_table%>%.[atom.id==n]%>%.[,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.2=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1.data_table%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.3=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n]%>%.[,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.4=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,times = 200
)
```


Compare using different dataset and different search methods. 

Results the data from fread is about the same as data converted from read.table. 

One step search is definitely to be faster than two step search. 



### Try to use setkey to see if there is an speed-up

First set the key for the entire dataset.

```{r}
atom.600.1_fread_withkey=atom.600.1_fread
setkey(atom.600.1_fread_withkey,atom.id)
```

Then compare the one that with key and one without the key.


```{r}
microbenchmark(
   
  MSD.all.matrix.600.nokey=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.withkey=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD
  }
  ,times = 200
)
```

Using key give a slightly worse performance but I would say they are statistically about the same. No data so that no t.test could be done.

### Let us look at the MSD function to see if we could improve its speed.

Define the MSD in a data.table way. 


```{r}
MSD2=function(data){
    MSD=(data[,xu]-data[1,xu])^2+(data[,yu]-data[1,yu])^2+(data[,zu]-data[1,zu])^2
    }
```


```{r}
microbenchmark(
   
  MSD.all.matrix.600.nokey_MSD1=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.withkey_MSD1=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.nokey_MSD2=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD2
  }
  ,
  MSD.all.matrix.600.withkey_MSD2=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD2
  }
  ,times = 200
)
```

The original MSD function gives better results than the new MSD2 function. About double the speed. Plus, using key gives slightly better performance. very small improvement.


Next, we need to compare if the calculation with MSD and without MSD in order to see if MSD calculation takes a long time. 

```{r}
microbenchmark(
   
  MSD.all.matrix.600.nokey_MSD1=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.nokey_noMSD=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]
  }
  ,
  MSD.all.matrix.600.withkey_MSD1=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD
  }
  ,
   MSD.all.matrix.600.withkey_noMSD=foreach(n=1:20,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]
  }
  ,times = 200
)
```


About a 3 millisecond more for calculation with MSD. Possibly could improve with C code. But overall effect would be small?
Plus, using key gives a marginally quicker calculation. 

### Next, let us see if the foreach function takes a long time 

Need to know which MSD and data do you use. 

Use the original MSD function, and the data from fread with and without key.

```{r}
microbenchmark(
  MSD.all.matrix.600.nokey_MSD1_noforeach=atom.600.1_fread%>%.[atom.id==100,.(xu,yu,zu)]%>%MSD
  ,
  MSD.all.matrix.600.withkey_MSD1_noforeach=atom.600.1_fread_withkey%>%.[.(100),.(xu,yu,zu)]%>%MSD
  ,
   MSD.all.matrix.600.nokey_MSD1=foreach(n=1:100,.combine=rbind)%do%{ 
    atom.600.1_fread%>%.[atom.id==n,.(xu,yu,zu)]%>%MSD
  }
  ,
  MSD.all.matrix.600.withkey_MSD1=foreach(n=1:100,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD
  }
  ,times = 200
)
```

There is some room for improvement here for using foreach. Using foreach gives a much slower calculation. Like a lot slower.


### check if the various data.table methods give the same answer as the the dplyr


```{r}
# two step search
(  (atom.600.1.data_table%>%.[atom.id==2000]%>%.[,.(xu,yu,zu)]%>%MSD)==(atom.600.1%>%filter(atom.id==2000)%>%select(xu,yu,zu)%>%MSD) )%>%sum()
# one step search
(  (atom.600.1_fread%>%.[atom.id==2000,.(xu,yu,zu)]%>%MSD)==(atom.600.1%>%filter(atom.id==2000)%>%select(xu,yu,zu)%>%MSD) )%>%sum()

# one step search with key
(  (atom.600.1_fread_withkey%>%.[atom.id==2000,.(xu,yu,zu)]%>%MSD)==(atom.600.1%>%filter(atom.id==2000)%>%select(xu,yu,zu)%>%MSD) )%>%sum()
# one step search with key and new MSD function
(  (atom.600.1_fread_withkey%>%.[atom.id==2000,.(xu,yu,zu)]%>%MSD2)==(atom.600.1%>%filter(atom.id==2000)%>%select(xu,yu,zu)%>%MSD) )%>%sum()


# one step search with key
(  (atom.600.1_fread_withkey%>%.[.(2000),.(xu,yu,zu)]%>%MSD)==(atom.600.1%>%filter(atom.id==2000)%>%select(xu,yu,zu)%>%MSD) )%>%sum()
# one step search with key and new MSD function
(  (atom.600.1_fread_withkey%>%.[.(2000),.(xu,yu,zu)]%>%MSD2)==(atom.600.1%>%filter(atom.id==2000)%>%select(xu,yu,zu)%>%MSD) )%>%sum()
```



It seems that using purely the key is faster but gives different result.

Find out the difference between using key and not using key 


```{r}
((atom.600.1_fread_withkey%>%.[atom.id==1000,.(xu,yu,zu)]%>%MSD2)-(atom.600.1_fread_withkey%>%.[.(1000),.(xu,yu,zu)]%>%MSD) )%>%plot()
```

The difference is kind of big, but why? Read the cheat sheet for data.table. But leave it for now, since this is not a serious bottleneck right now. 


```{r}
setkey(atom.600.1_fread_withkey,atom.id)

(atom.600.1_fread_withkey%>%.[atom.id==1000,.(xu,yu,zu)])%>%dim()
(atom.600.1_fread_withkey[.(1000),.(xu,yu,zu)])%>%dim()

```

Using key to search for the value is not doing as it is stated in the cheat-sheet from datacamp for data.table package. It only returned one value. 

Read more about it later. I figured out how to work with the key. If the key column is integer or numeric, you will need to do .(10) to select key rows that have a value of 10. 



After much test, the possible improvement could be done with the foreach, and possible improvement with MSD written in C. 

Let us try to optimize on how to combine the results and compare it with foreach. 

```{r}
## Need to define MSD function and load data.table library before using this function. Plus the the key for 
MSD_matrix=function(data,timestep,tot_atoms){
  MSD_empty_matrix=matrix(NA,nrow=tot_atoms,ncol=timestep)
  for(i in 1:tot_atoms){
    MSD_empty_matrix[i,]=data%>%.[.(i),.(xu,yu,zu)]%>%MSD
  }
  return(MSD_empty_matrix)
}
```




```{r}
microbenchmark(
  
  MSD.all.matrix.600.withkey_MSD1_for=MSD_matrix(data=atom.600.1_fread_withkey,timestep = 5001,tot_atoms = 10)
  ,
  
   MSD.all.matrix.600.withkey_MSD1_foreach=foreach(n=1:10,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD
   }
  ,times = 100
  
)
```

Using for within a function with replacement is the fastest method so far. The improvement is not dramatic 



check if they gives the same results.

```{r}

(  (MSD_matrix(data=atom.600.1_fread_withkey,timestep = 5001,tot_atoms = 1))==(atom.600.1%>%filter(atom.id==1)%>%select(xu,yu,zu)%>%MSD) )%>%sum()
```

check if they gives the same results.

```{r}
(
  MSD_matrix(data=atom.600.1_fread_withkey,timestep = 5001,tot_atoms = 10)[5,]==(foreach(n=1:100,.combine=rbind)%do%{ 
    atom.600.1_fread_withkey%>%.[.(n),.(xu,yu,zu)]%>%MSD
   })[5,]
 )%>%sum
```



























































