---
title: 'Predict glass transition from short time simulation '
author: "Zijun Lu"
output:
  html_document: default
  
---

## recap on what we are trying to achieve
We are tryng to use a **short time (~14ps) simulation** to get the glass transition temperature compatible with long time simulation.

```{r,echo=FALSE,include=FALSE}

library(foreach)
library(dplyr)
Data.import=function(Path="~/Dropbox/lammps" , filename = "MSD.colmean.matrix.1.txt", temp=c(320,345,370,395,420,445,470,500,550,600),polymer="PMMA"){
  
  # read in data into a matrix
 MSDcol=foreach(i=1:length(temp),.combine = cbind)%do% {
   # set correct path for the data file
   path=paste(Path,"/", polymer,"/atom",temp[i], sep='')
   path
   setwd(path)
   # read in the data
   MSDcol=read.table(file=filename,header=TRUE,sep=',')
   library(magrittr)
   MSDcol%>%dim
   MSDcol=MSDcol[[1]]
   return(MSDcol)
 }
 
 # add a time variable 
 time=1:dim(MSDcol)[1]
 # change data into data frame 
  MSDcol=cbind(time,MSDcol)%>%as.data.frame()
  
}

```

## MSD g0 plot for PS

```{r,echo=FALSE}

#Run (optimizaed data import function.R) file first 
# temp=c(200,250,300,350,400,450,500,550,600)
# import g0 MSD colmean data for PS 
########################################################
library(dplyr)

MSD.PS.g0=Data.import(filename="MSD.colmean.matrix.1.txt", temp=c(200,250,300,350,400,450,500,550,600),polymer="PS")

MSD.PS.g0=MSD.PS.g0%>%as.data.frame()

# add correct colname 
temp=c(200,250,300,350,400,450,500,550,600)
colnames(MSD.PS.g0)=c("time",paste("T",temp,sep=""))

#regular plot 
###########################################
library(reshape2)
MSD.PS.g0.melt=melt(MSD.PS.g0,id.vars="time")
library(dplyr)

library(ggplot2)
ggplot(data=MSD.PS.g0.melt,aes(x=time,y=value,colour=variable)) +geom_point(size=0.5)+
  ylab("MSD")+
  ggtitle("PS g0 MSD vs time plot")

# log-log plot 
###########################################
library(reshape2)
MSD.PS.g0.melt=melt(MSD.PS.g0,id.vars="time")


library(ggplot2)
ggplot(data=MSD.PS.g0.melt,aes(x=log(time),y=log(value),colour=variable)) +geom_point(size=0.5)+
  ylab("log(MSD)")+
  ggtitle("PS g0 MSD vs time log-log plot")


```

## MSD g0 plot for PMMA



```{r}
library(dplyr)

MSD.PMMA.g0=Data.import(filename="MSD.colmean.matrix.1.txt", temp=c(320,345,370,395,420,445,470,500,550,600),polymer="PMMA")

MSD.PMMA.g0=MSD.PMMA.g0%>%as.data.frame()

# add correct colname 
temp=c(320,345,370,395,420,445,470,500,550,600)
colnames(MSD.PMMA.g0)=c("time",paste("T",temp,sep=""))

#regular plot 
###########################################
library(reshape2)
MSD.PMMA.g0.melt=melt(MSD.PMMA.g0,id.vars="time")
library(dplyr)

library(ggplot2)
ggplot(data=MSD.PMMA.g0.melt,aes(x=time,y=value,colour=variable)) +geom_point(size=0.5)+
  ylab("MSD")+
  ggtitle("PMMA g0 MSD vs time plot")

# log-log plot 
###########################################
library(reshape2)
MSD.PMMA.g0.melt=melt(MSD.PMMA.g0,id.vars="time")


library(ggplot2)
ggplot(data=MSD.PMMA.g0.melt,aes(x=log(time),y=log(value),colour=variable)) +geom_point(size=0.5)+
  ylab("log(MSD)")+
  ggtitle("PMMA g0 MSD vs time log-log plot")
```

## MSD g0 plot for PIB



```{r}
library(dplyr)

MSD.PIB.g0=Data.import(filename="MSD.colmean.matrix.1.txt", temp=c(150,175,200,225,250,275,300,350,400),polymer="PIB")

MSD.PIB.g0=MSD.PIB.g0%>%as.data.frame()

# add correct colname 
temp=c(150,175,200,225,250,275,300,350,400)
colnames(MSD.PIB.g0)=c("time",paste("T",temp,sep=""))

#regular plot 
###########################################
library(reshape2)
MSD.PIB.g0.melt=melt(MSD.PIB.g0,id.vars="time")
library(dplyr)

library(ggplot2)
ggplot(data=MSD.PIB.g0.melt,aes(x=time,y=value,colour=variable)) +geom_point(size=0.5)+
  ylab("MSD")+
  ggtitle("PIB g0 MSD vs time plot")

# log-log plot 
###########################################
library(reshape2)
MSD.PIB.g0.melt=melt(MSD.PIB.g0,id.vars="time")


library(ggplot2)
ggplot(data=MSD.PIB.g0.melt,aes(x=log(time),y=log(value),colour=variable)) +geom_point(size=0.5)+
  ylab("log(MSD)")+
  ggtitle("PIB g0 MSD vs time log-log plot")
```

## MSD g2 plot for PS

```{r,echo=FALSE}

#Run (optimizaed data import function.R) file first 
# temp=c(200,250,300,350,400,450,500,550,600)
# import g2 MSD colmean data for PS 
########################################################
library(dplyr)

MSD.PS.g2=Data.import(filename="MSD.COM.colmean.1.txt", temp=c(200,250,300,350,400,450,500,550,600),polymer="PS")

MSD.PS.g2=MSD.PS.g2%>%as.data.frame()

# add correct colname 
temp=c(200,250,300,350,400,450,500,550,600)
colnames(MSD.PS.g2)=c("time",paste("T",temp,sep=""))

#regular plot 
###########################################
library(reshape2)
MSD.PS.g2.melt=melt(MSD.PS.g2,id.vars="time")
library(dplyr)

library(ggplot2)
ggplot(data=MSD.PS.g2.melt,aes(x=time,y=value,colour=variable)) +geom_point(size=0.5)+
  ylab("MSD")+
  ggtitle("PS g2 MSD vs time plot")

# log-log plot 
###########################################
library(reshape2)
MSD.PS.g2.melt=melt(MSD.PS.g2,id.vars="time")


library(ggplot2)
ggplot(data=MSD.PS.g2.melt,aes(x=log(time),y=log(value),colour=variable)) +geom_point(size=0.5)+
  ylab("log(MSD)")+
  ggtitle("PS g2 MSD vs time log-log plot")


```

## MSD g2 plot for PMMA



```{r}
library(dplyr)

MSD.PMMA.g2=Data.import(filename="MSD.COM.colmean.1.txt", temp=c(320,345,370,395,420,445,470,500,550,600),polymer="PMMA")

MSD.PMMA.g2=MSD.PMMA.g2%>%as.data.frame()

# add correct colname 
temp=c(320,345,370,395,420,445,470,500,550,600)
colnames(MSD.PMMA.g2)=c("time",paste("T",temp,sep=""))

#regular plot 
###########################################
library(reshape2)
MSD.PMMA.g2.melt=melt(MSD.PMMA.g2,id.vars="time")
library(dplyr)

library(ggplot2)
ggplot(data=MSD.PMMA.g2.melt,aes(x=time,y=value,colour=variable)) +geom_point(size=0.5)+
  ylab("MSD")+
  ggtitle("PMMA g2 MSD vs time plot")

# log-log plot 
###########################################
library(reshape2)
MSD.PMMA.g2.melt=melt(MSD.PMMA.g2,id.vars="time")


library(ggplot2)
ggplot(data=MSD.PMMA.g2.melt,aes(x=log(time),y=log(value),colour=variable)) +geom_point(size=0.5)+
  ylab("log(MSD)")+
  ggtitle("PMMA g2 MSD vs time log-log plot")
```

## MSD g2 plot for PIB

Need to calculate the g2 MSD



## Method 1: Taylor expansion of MSD with Brownian motion approximation $\langle x^2(t) \rangle =2Dt$  .



Let us first define the velocity auto-correlation function for a stationary process **x(t)**
$$
    C_{\upsilon}(t_2,t_1)=<\upsilon(t_2),\upsilon(t_1)>=\int\int d\upsilon_1d\upsilon_2P(\upsilon_2,t_2|\upsilon_1,t_1) W(\upsilon_1,t_1)\upsilon_1\upsilon_2
$$
$P(\upsilon_2,t_1|\upsilon_1,t_1)$ is the transition probability density from velocity $\upsilon_1$ at $t_1$ to velocity $\upsilon_2$ at $t_2$. 

$W(\upsilon_1,t_1)$ is the probability density finding particle with velocity $\upsilon_1$ at $t_1$

### Green Kubo relation 

Thus, the stochastic process x(t) has constant mean and constant variance, which means that $C_{\upsilon}(t_2,t_1)=C_{\upsilon}(t_2-t_1,0)=C_{\upsilon}(\tau,0)$

Recall that $$<x(t)x(t)>=\int^t_0 \int^t_0 dt_1dt_2 \times<\upsilon(t_2),\upsilon(t_1)>=\int^t_0\int^t_0 dt_1dt_2 C_{\upsilon}(t_2,t_1)$$ 


Thus, $$D=\frac{<x^2(t)> }{2t} =\int^t_0 C_{\upsilon}(\tau,0)d\tau$$

Taylor expansion of $C_{\upsilon}(t_2,t_1)$ is


\begin{align*}
    C_{\upsilon}(\tau+t,t)&=<\upsilon(\tau+t),\upsilon(t)>=\sum\limits_{n=0}^{\infty} \frac{\tau^{2n}}{2n!} <\upsilon^{(2n)}(t),\upsilon(t)>\\
    &\textit{Since that} <\upsilon^{(2)}(t),\upsilon(t)>=-<\upsilon^{(1)}(t),\upsilon^{(1)}(t)>\\
    &=\sum\limits_{n=0}^{\infty}(-1)^{2n} \frac{\tau^{2n}}{2n!} <\upsilon^{(n)}(t),\upsilon^{(n)}(t)>\\
    &=(<\upsilon(t),\upsilon(t)>-\frac{1}{2}\alpha \tau^2+\frac{1}{4}\beta \tau^4+O(\tau^6))
\end{align*}




where $\alpha=<\upsilon^{(1)}(t),\upsilon^{(1)}(t)>$.

### Expression for alpha  

\begin{align*}
    <x^2(t)>=&2t\int^t_0 C_{\upsilon}(\tau,0)d\tau\\
    &=2t\int^t_0d\tau (<\upsilon(t),\upsilon(t)>-\frac{1}{2}\alpha \tau^2+O(\tau^4))\\  
    &\textit{Since that }<\upsilon(t),\upsilon(t)>=\frac{3k_BT}{m}\\  
    &=2t^2(\frac{3k_BT}{m}-\frac{1}{6}\alpha t^2 +O(t^5))
\end{align*}

\begin{align*}
    \frac{d^2}{dt^2}[\frac{<x^2(t)>}{t^2}]\simeq-&\frac{1}{3}\alpha=-\frac{1}{3}<\upsilon^{(1)}(t),\upsilon^{(1)}(t)>\\&=-\frac{1}{3m^2}<F(t),F(t)>
\end{align*}

**In actual calculation, Solomon's method is to use linear regression with all even orders on MSD and extract the coefficient for t^4.**

## alpha term : MSD by averaging over all atoms for PMMA, PS and PIB.  

First, change the units of time and distance to ps and nm. The original timestep is 10fs, and distance unit is angstrom. So I achieve it by dividing MSD by 100 and (1:tot_num_of_timestep)*0.01. 

Second, the linear regression is done as $MSD/t^2=c_0+c_1*t^2+c_2*t^4+...+c_5*t^{10}$. Then the alpha term (curvature of MSD) is extracted from the simulation by extracting $c_2$ from the linear regression coeeficients. 


```{r,echo=FALSE}
# for even term in t alpha and NGP plotting
###############################################################################################################
#Run ( data_import.R) file first 




# write a function to plot the alpha term for different polymers and different MSD calculation method. 
#Run (optimizaed data import function.R) file first 


alpha.even.ngp.plot=function(filename1="MSD.colmean.matrix.1.txt",filename2="NGP.1.txt", temp=c(200,250,300,350,400,450,500,550,600),polymer="PS"){
  
  
  # import colmean data and do linear regression for alpha term 
  ##############################################################
  MSD.PS.g0=Data.import(filename=filename1, temp=temp,polymer=polymer)
  
  MSD.PS.g0=MSD.PS.g0%>%as.data.frame()
  
  
  
  # Regression method 1 : using all even order and no intercept
  ################################################################
  
  # change the time units to ps
  time=MSD.PS.g0[,1]*0.01
  # change  the length unit to nm. 
  MSD.PS.g0=MSD.PS.g0/100
  library(foreach)
  
  alpha=foreach(i=2:(dim(MSD.PS.g0)[2]),.combine = rbind)%do%{
    (lm(MSD.PS.g0[,i]/time^2~I(time^2)+I(time^4)+I(time^6)+I(time^8)+I(time^10))%>%confint(level=0.95))[2,]
  }
  Rsquared=foreach(i=2:(dim(MSD.PS.g0)[2]),.combine = rbind)%do%{
     c(  ( lm(MSD.PS.g0[,i]/time^2~I(time^2)+I(time^4)+I(time^6)+I(time^8)+I(time^10)) )%>%summary%>%.$adj.rsquared,
          ( lm(MSD.PS.g0[,i]/time^2~I(time^2)+I(time^4)+I(time^6)+I(time^8)+I(time^10)) )%>%summary%>%.$r.squared
        )
  }
  
  ###############################################################
  
  #first transform the matrix then turn it into a vector 
  alpha=alpha%>%t%>%as.vector
  # combine alpha and temp into a data frame 
  temp=rep(temp,each=2)
  alpha=data.frame(alpha,temp,Rsquared,row.names = NULL)
  
  
  
  # calculate the average between two alpha values 
  alpha_average=alpha%>%group_by(temp) %>% summarize(alpha_average=mean(alpha))
  alpha_average=rbind(alpha_average,alpha_average)%>%arrange(temp)
  
  alpha=cbind(alpha,alpha_average["alpha_average"])
  
  
  # unscaled plot for two 95% confidence interval and average of the end points of the range 
  library(ggplot2)
  
  p1=ggplot(data=alpha, aes(y=alpha/temp,x=temp))+geom_point()+geom_line(aes(x=temp,y=alpha_average/temp),color="red")+
    #geom_smooth(method= "loess")+
    xlab("Temperature")+ylab("Curvature of MSD/Temperature ")
  p3=ggplot(data=alpha, aes(y=Rsquared,x=temp))+geom_point(color="orange")+
    xlab("Temperature")+ylab("R^2 and adj R^2")
  ####################################################################################################################
  
  #NGP 
  
  NGP=Data.import(filename=filename2, temp=temp,polymer=polymer)
  
  
  temp=temp
  # add column names
  colnames(NGP)=c("time",paste("T",temp,sep=""))
  # delete NA values
  NGP=NGP[complete.cases(NGP),]
  
  NGP=NGP%>%colMeans()
  
  NGP=NGP[-1]
  
  
  
  
  NGP=data.frame(NGP,temp)
  library(ggplot2)
  p2=ggplot(data=NGP, aes(y=NGP,x=temp))+geom_line()+
    xlab("Temperature")+ylab("NGP value")
    
  
  
  ## Plot NGP and alpha term together
  
  
  library(grid)
  grid.newpage()
  grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(p3), size = "first"))
  
  
  
  
  
  
}  


# for all terms in t alpha and NGP plotting
#################################################################################################################

```



## PS

### PS g0 alpha term all even orders (t^2 to t^20)

```{r}
alpha.even.ngp.plot(filename1="MSD.colmean.matrix.1.txt",filename2="NGP.1.txt", temp=c(200,250,300,350,400,450,500,550,600),polymer="PS")

```

## PMMA 

### PMMA g0 alpha term even orders (t^2 to t^20)

```{r}
alpha.even.ngp.plot(filename1="MSD.colmean.matrix.1.txt",filename2="NGP.1.txt", temp=c(320,345,370,395,420,445,470,500,550,600),polymer="PMMA")

```

## PIB 

### PIB g0 alpha term even orders (t^2 to t^20)

```{r}
alpha.even.ngp.plot(filename1="MSD.colmean.matrix.1.txt",filename2="NGP.1.txt", temp=c(150,175,200,225,250,275,300,350,400),polymer="PIB")

```



## PS

### PS g2 alpha term all even orders (t^2 to t^20)

```{r}
alpha.even.ngp.plot(filename1="MSD.COM.colmean.1.txt",filename2="NGP.1.txt", temp=c(200,250,300,350,400,450,500,550,600),polymer="PS")

```

## PMMA 

### PMMA g2 alpha term even orders (t^2 to t^20)

```{r}
alpha.even.ngp.plot(filename1="MSD.COM.colmean.1.txt",filename2="NGP.1.txt", temp=c(320,345,370,395,420,445,470,500,550,600),polymer="PMMA")

```

## PIB 

### PIB g2 alpha term even orders (t^2 to t^20)

Need to calculate.


## alpha term : MSD by averaging over same atom type 

## PS

### atom type all even orders+Non_gaussian parameter (t^2 to t^20)

```{r,echo=FALSE}



#Run (optimizaed data import function.R) file first 

# write a function to import data, analyze and print out the graph. 
########################################################################################################################
atom.type.plot=function(temp, polymer, type){
  filename1=paste("atom.type.",type,".MSD.colmean.txt",sep="")
  filename2=paste("atom.type.",type,".NGP.txt",sep="")
  # alpha term
  #############################################################################
  # import MSD colmean data
  ########################################################
  MSD=Data.import(filename=filename1, temp=temp,polymer=polymer)
  
  MSD=MSD%>%as.data.frame()
  
  
  
  # Regression method 1 : using all even order and no intercept
  ########################################################
  
  
  
  
  
  
  time=MSD[,1]*0.01
  MSD=MSD/100
  
  library(foreach)
  
  alpha=foreach(i=2:(dim(MSD)[2]),.combine = rbind)%do%{
    (lm(MSD[,i]/time^2~I(time^2)+I(time^4)+I(time^6)+I(time^8)+I(time^10))%>%confint(level=0.95))[2,]
  }
  
  Rsquared=foreach(i=2:(dim(MSD)[2]),.combine = rbind)%do%{
     c(  ( lm(MSD[,i]/time^2~I(time^2)+I(time^4)+I(time^6)+I(time^8)+I(time^10)) )%>%summary%>%.$adj.rsquared,
          ( lm(MSD[,i]/time^2~I(time^2)+I(time^4)+I(time^6)+I(time^8)+I(time^10)) )%>%summary%>%.$r.squared
        )
  }
  #first transform the matrix then turn it into a vector 
  alpha=alpha%>%t%>%as.vector
  # combine alpha and temp into a data frame 
  temp=rep(temp,each=2)
  alpha=data.frame(alpha,temp,Rsquared,row.names = NULL)
  
  
  
  
  # calculate the average between two alpha values 
  alpha_average=alpha%>%group_by(temp) %>% summarize(alpha_average=mean(alpha))
  alpha_average=rbind(alpha_average,alpha_average)%>%arrange(temp)
  
  alpha=cbind(alpha,alpha_average["alpha_average"])
  
  alpha
  
  #########################################################################################
  alpha.title1=paste("alpha term for", polymer, "atom type", type, sep=" ")
  # unscaled plot 
  library(ggplot2)
  
  p1=ggplot(data=alpha, aes(y=alpha/temp,x=temp))+geom_point()+geom_line(aes(x=temp,y=alpha_average/temp))+
    xlab("Temperature")+ylab("Curvature of MSD/Temperature ")+ggtitle(alpha.title1)
  p3=ggplot(data=alpha, aes(y=Rsquared,x=temp))+geom_point(color="orange")+
    xlab("Temperature")+ylab("R^2 and adj R^2")
  #NGP
  ##################################################################################################
  #NGP 
  
  NGP=Data.import(filename=filename2, temp=temp,polymer=polymer)
  
  
  temp=temp
  # add column names
  colnames(NGP)=c("time",paste("T",temp,sep=""))
  # delete NA values
  NGP=NGP[complete.cases(NGP),]
  
  NGP=NGP%>%colMeans()
  
  NGP=NGP[-1]
  
  NGP
  
  NGP.title=paste("NGP for", polymer, "atom type", type,sep=" ")
  NGP=data.frame(NGP,temp)
  library(ggplot2)
  p2=ggplot(data=NGP, aes(y=NGP,x=temp))+geom_line()+
    xlab("Temperature")+ylab("NGP value")+
    ggtitle(NGP.title)
  
  
  ## Plot NGP and alpha term together
  
  
  library(grid)
  grid.newpage()
  grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2),ggplotGrob(p3), size = "first"))
  
}
##########################################################################################################




#PMMA
##########################################################################################################
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=1)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=2)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=3)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=4)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=5)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=6)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=7)
atom.type.plot(temp=c(320,345,370,395,420,445,470,500,550,600), polymer="PMMA", type=8)


# PS
##########################################################################################################
atom.type.plot(temp=c(200,250,300,350,400,450,500,550,600), polymer="PS", type=1)
atom.type.plot(temp=c(200,250,300,350,400,450,500,550,600), polymer="PS", type=2)
atom.type.plot(temp=c(200,250,300,350,400,450,500,550,600), polymer="PS", type=3)
atom.type.plot(temp=c(200,250,300,350,400,450,500,550,600), polymer="PS", type=4)
atom.type.plot(temp=c(200,250,300,350,400,450,500,550,600), polymer="PS", type=5)
atom.type.plot(temp=c(200,250,300,350,400,450,500,550,600), polymer="PS", type=6)
# PIB
##########################################################################################################
atom.type.plot(temp=c(175,200,225,250,275,300,350,400), polymer="PIB", type=1)
atom.type.plot(temp=c(175,200,225,250,275,300,350,400), polymer="PIB", type=2)
atom.type.plot(temp=c(175,200,225,250,275,300,350,400), polymer="PIB", type=3)
atom.type.plot(temp=c(175,200,225,250,275,300,350,400), polymer="PIB", type=4)
atom.type.plot(temp=c(175,200,225,250,275,300,350,400), polymer="PIB", type=5)

```



